---
title: "Spatial Data"
output: github_document
---

*Note:* The main goal for this exercise is to practice data visualization for spatial data, no data wrangling or any other phase in the process, so the code is going to be kept to the minumum and focus to the goal, only where it makes sense I will clean up the data.

Setting up environment
```{r}
library(tidyverse)

countries_geodata <- read.csv("../data/CountryMapCoordinates.csv")
suicide_rates_dataset <- read.csv("../data/SuicideRatesOverview1985-to-2016.csv")
suicide_dataset <- suicide_rates_dataset %>%
  merge(countries_geodata, by.x = "country", by.y = "name")

#countries_geodata <- NULL
#suicide_rates_dataset <- NULL

```

# Plots

Generating base GGPlot Plot

```{r}
myplot <- ggplot() +
  xlab("") +
  ylab("") +
  theme(
    panel.background = element_blank(),
    axis.title.x =element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y =element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())
  
myplot +
  ggtitle("Base Map of the world!") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90")
  
```

##Dot Density Map

Let's see where we have suicide records, keep in mind this is a sample.

```{r}
myplot +
  ggtitle("Suicide by Countries (1985 - 2016)") +
  borders(
    database = "world",
    colour = "gray60",
    fill = "grey90") +
  geom_point(
    data = suicide_dataset,
    aes(
      x = longitude,
      y = latitude))
```

## Contour Map
Let's see how it is concentrated.

```{r}
myplot +
  ggtitle("Density of Suicides by Country") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "gray90") +
  geom_density2d(
    data = suicide_dataset,
    aes(
     x = longitude,
     y = latitude))
```

Let's zoom into Europe

```{r}
myplot +
  ggtitle("Density of Suicides in Europe") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  geom_density2d(
    data = suicide_dataset,
    aes(
      x = longitude,
      y = latitude)) +
  coord_cartesian(
    xlim = c(-20,59),
    ylim = c(35,71))
```

Let's Zoom into Latin America

```{r}
myplot +
  ggtitle("Density of Suicide in Latin America") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  geom_density2d(
    data = suicide_dataset,
    aes(
      x = longitude,
      y = latitude)) +
  coord_cartesian(
    xlim = c(-130,-35),
    ylim = c(30,-50))
  
```

## Level Map

```{r}
options(scipen=999)
myplot +
  ggtitle("Density of Suicide by Country (1985-2016)") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  stat_density2d(
    data = suicide_dataset,
    aes(
      x = longitude,
      y = latitude,
      alpha = ..level..),
    geom = "polygon",
    fill = "blue") +
  labs(alpha = "Density")
```

## Bubble Map

I'm only displaying the total over the time, from 1985 and 2016, but we should pay attention to the proportion based on the total of people by country, so we may find countries with more total suicides but having larger population the proportion may be lower... We also may find some countries have less suicides ver time (and viceversa), I will display those examples in further exercises.

```{r}

suicide_grouped_dataset <- suicide_dataset %>%
  select(country, suicides_no,longitude, latitude) %>%
  group_by(country, longitude, latitude) %>%
  summarize(Count = sum(suicides_no))

suicide_grouped_dataset %>%
  ggplot() +
  xlab("") +
  ylab("") +
  theme(
    panel.background = element_blank(),
    axis.title.x =element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y =element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank())+
  ggtitle("Count of Suicides by Country (1985-2016)") +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  geom_point(
    aes(
      x = longitude,
      y = latitude,
      size = Count)) +
  labs(size = "Suicides")
```

## Map Data

```{r}
map <- map_data("world")
head(map)
```

```{r}
suicide_dataset <- suicide_rates_dataset %>%
  left_join(map, by =c("country" = "region")) %>%
  select(
    country,
    Longitude = long,
    Latitude = lat,
    Group = group,
    Order = order,
    Suicides100k = suicides.100k.pop) %>%
  arrange(Order) %>%
  as.data.frame()
```

## Choropleth

```{r}
ggplot(data = suicide_dataset) +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  xlab("") +
  ylab("") +
  theme( panel.background = element_blank(),
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.title.y=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  geom_polygon(
    aes(
      x = Longitude,
      y = Latitude,
      group = Group,
      fill = Suicides100k),
    colour = "grey60") +
  scale_fill_gradient(
    low = "white",
    high = "red") +
  ggtitle("Count of Suicides by Country") +
  labs(color = "Suicides")
  
```

```{r}
ggplot(data = suicide_dataset) +
  borders(
    database = "world",
    colour = "grey60",
    fill = "grey90") +
  xlab("") +
  ylab("") +
  theme( panel.background = element_blank(),
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.title.y=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  coord_map(
    projection = "ortho",
    orientation = c(41,-74,0)) +
  geom_polygon(
    aes(
     x = Longitude ,
     y = Latitude,
     group = Group,
     fill = Suicides100k),
    colour = "grey60") +
  scale_fill_gradient(
    low = "white",
    high = "red") +
  ggtitle("Count of Suicides by Country") +
  labs(color = "Suicides")
```

