---
title: "Mastering Data Visualization - Text Data"
output: github_document
---

# Visualizing Text Data

## Text Data Analysis

Setting up environment

```{r}

library(tm) # text mining package
library(SnowballC)
library(wordcloud)
plots <- read.csv("../data/Plots.csv")
plots$Plot <- as.character(plots$Plot)
head(plots$Plot, 3)

```

Convert plots into corpus
```{r}
corpus <- Corpus(VectorSource(plots$Plot))
corpus[[1]]$content
```
Convert text to lower case
```{r}
corpus <- tm_map(corpus, content_transformer(tolower))
```

Remove punctuation
```{r}
corpus <- tm_map(corpus, removePunctuation)
```

Remove stop words from corpus
```{r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
```

Reduce terms to stems in corpus
```{r}
corpus <- tm_map(corpus, stemDocument, "english")
```


Strip whitespace from corpus
```{r}
corpus <- tm_map(corpus, stripWhitespace)
```

Convert corpus to text document
```{r}
corpus <- tm_map(corpus, PlainTextDocument)
```
 
Taking a look at the current result
```{r}
corpus[[1]]$content
```


Create a frequency word cloud
```{r}
wordcloud(
    words = corpus,
    max.words = 50)
```


# Charts/Plots


## Frequency Word Cloud

```{r}

```


## Quantitative Word Cloud

```{r}

```

    
## Colored Word Cloud

```{r}

```

