---
title: "Mastering Data Visualization - Text Data"
output: github_document
---

# Visualizing Text Data

## Text Data Analysis

Setting up environment

```{r}

library(tm) # text mining package
library(SnowballC)
library(wordcloud)
library(RColorBrewer)
plots <- read.csv("../data/Plots.csv")
plots$Plot <- as.character(plots$Plot)
head(plots$Plot, 3)

```

Convert plots into corpus
```{r}
corpus <- Corpus(VectorSource(plots$Plot))
corpus[[1]]$content
```
Convert text to lower case
```{r}
corpus <- tm_map(corpus, content_transformer(tolower))
```

Remove punctuation
```{r}
corpus <- tm_map(corpus, removePunctuation)
```

Remove stop words from corpus
```{r}
corpus <- tm_map(corpus, removeWords, stopwords("english"))
```

Reduce terms to stems in corpus
```{r}
corpus <- tm_map(corpus, stemDocument, "english")
```


Strip whitespace from corpus
```{r}
corpus <- tm_map(corpus, stripWhitespace)
```

Convert corpus to text document
```{r}
corpus <- tm_map(corpus, PlainTextDocument)
```
 
Taking a look at the current result
```{r}
head(corpus[[1]]$content)
```

# Charts/Plots


## Frequency Word Cloud

Create a frequency word cloud
```{r}
wordcloud(
    words = corpus[1],
    max.words = 50)
```



## Quantitative Word Cloud

Setting up environtment
```{r}
words <- read.csv("../data/Words.csv")
head(words, n=10)
```


```{r}
wordcloud(
  words = words$Term,
  freq = words$Box.Office,
  max.words = 50,
  scale = c(2,0.1))
```

    
## Colored Word Cloud

```{r}
palette <- brewer.pal(
  n = 9,
  name = "Oranges")

colors <- palette[cut(words$Critic.Score, 9)]

wordcloud(
  words = words$Term,
  freq = words$Box.Office,
  max.words = 50,
  scale = c(2,0.5),
  colors = colors,
  ordered.colors = TRUE)
```

